Concurrent Programming with CXL

### 2020-06-29 book/paper.tex
Formatting up to monitors

An important invariant of the SBS technique is that the sum of the $N+1$ semaphores
must always be 0 or 1, and it should be 0 when a process is accessing the shared
variables and 1 when not.
As it were, these semaphores taken together are emulating a single binary semaphore.
Initially, \textit{mutex} is~1 and the other $N$ semaphores
are all~0.  To maintain the invariant, each process alternates \texttt{P} and
\texttt{V} operations, starting with \texttt{P(\&mutex)} when it tries to
enter the CCS, and ending with a \texttt{V} operation on one of the semaphores
(yet to be determined).

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \texttt{r\_entered}: the number of readers in the critical section;
\item \texttt{w\_entered}: the number of writers in the critical section;
\item \texttt{r\_waiting}: the number of readers waiting to enter;
\item \texttt{w\_waiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
To wit, check out \texttt{rlock\_acquire}.  It first checks to see if there
are any writers.  If so, it increments \texttt{r\_waiting}.  If not,
it increments \texttt{r\_entered}.

@@

Let us return now to \texttt{rlock\_acquire}. In the case that the process,
say $p$,
found there is a writer in the critical section, it vacates
\textit{mutex} and procures \texttt{r\_sema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
$p$ by incrementing \texttt{r\_sema}, $p$ updates the state by decrementing
\texttt{r\_waiting} and incrementing \texttt{r\_entered}.  $p$
then invokes \texttt{V\_one()} one more time.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \textit{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\texttt{w\_entered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \texttt{r\_sema}.
The sum of the semaphores is now 1, because \texttt{r\_sema} is 1.  This
guarantees that only a waiting reader, procuring \texttt{r\_sema}, can enter
the critical section.
\item When it does, the reader decrements \texttt{r\_waiting}
from 2 to 1, and increments \texttt{r\_entered} from 0 to 1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \texttt{r\_sema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \texttt{r\_waiting} from 1 to 0 and
increments \texttt{r\_entered} from 1 to 2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \textit{mutex}.
\end{enumerate}

@@

The \emph{Alternating Bit Protocol} is suitable for our purposes.
We assume that there are two unreliable network channels: one from Alice
to Bob and one from Bob to Alice.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \texttt{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).  Let \texttt{s\_chan} be the channel
from Alice to Bob and \texttt{r\_chan} the channel from Bob to Alice.
\texttt{net\_send(pchan, m, reliable)} models sending a message \texttt{m} to
\texttt{\^{}pchan}, where \texttt{pchan} is either \texttt{\&s\_chan}
or \texttt{\&r\_chan}.
The method places either \texttt{m} (to model a successful send)
or \texttt{()} (to model loss) in \texttt{\^{}pchan}.
The use of the \texttt{reliable} flag will be explained later.
\texttt{net\_recv(pchan)} models checking \texttt{\^{}pchan} for
the next message.

>>>>

An important invariant of the SBS technique is that the sum of the $N+1$ semaphores
must always be 0 or 1, and it should be 0 when a process is accessing the shared
variables and 1 when not.
As it were, these semaphores taken together are emulating a single binary semaphore.
Initially, \textit{mutex} is~1 and the other $N$ semaphores
are all~0.  To maintain the invariant, each process alternates \texttt{P} and
\texttt{V} operations, starting with \texttt{P}(\&\textit{mutex}) when it tries to
enter the CCS, and ending with a \texttt{V} operation on one of the semaphores
(yet to be determined).

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \textit{r\_entered}: the number of readers in the critical section;
\item \textit{w\_entered}: the number of writers in the critical section;
\item \textit{r\_waiting}: the number of readers waiting to enter;
\item \textit{w\_waiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
To wit, check out \texttt{rlock\_acquire}.  It first checks to see if there
are any writers.  If so, it increments \textit{r\_waiting}.  If not,
it increments \textit{r\_entered}.

@@

Let us return now to \texttt{rlock\_acquire}. In the case that the process,
say $p$,
found there is a writer in the critical section, it vacates
\textit{mutex} and procures \textit{r\_sema},
blocking on the semaphore associated
with waiting readers.  When later some other process passes the baton to
$p$ by incrementing \textit{r\_sema}, $p$ updates the state by decrementing
\textit{r\_waiting} and incrementing \textit{r\_entered}.  $p$
then invokes \texttt{V\_one()} one more time.

As an example, consider the case where there is a writer in the critical
section and there are two readers waiting.  Let us see what happens when
the writer calls \texttt{wlock\_release}:
\begin{enumerate}
\item After procuring \textit{mutex},
the sum of the three semaphores is 0.  The writer then decrements
\textit{w\_entered} and calls \texttt{V\_one()}.
\item \texttt{V\_one()} finds that there are no writers in the critical section
and there are two readers waiting.  It therefore vacates \textit{r\_sema}.
The sum of the semaphores is now 1, because \textit{r\_sema} is 1.  This
guarantees that only a waiting reader, procuring \textit{r\_sema}, can enter
the critical section.
\item When it does, the reader decrements \textit{r\_waiting}
from 2 to 1, and increments \textit{r\_entered} from 0 to 1.
The reader finally calls \texttt{V\_one()}.
\item Again, \texttt{V\_one()} finds that there are no writers and
that there are readers waiting, so again it vacates \textit{r\_sema},
passing the baton to the one remaining waiting reader.
\item The remaining reader decrements \textit{r\_waiting} from 1 to 0 and
increments \textit{r\_entered} from 1 to 2.
\item Finally, the remaining reader  calls \texttt{V\_one()}.
\texttt{V\_one()} does not find any processes waiting,
and so it vacates \textit{mutex}.
\end{enumerate}

@@

The \emph{Alternating Bit Protocol} is suitable for our purposes.
We assume that there are two unreliable network channels: one from Alice
to Bob and one from Bob to Alice.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \textit{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).  Let \texttt{s\_chan} be the channel
from Alice to Bob and \textit{r\_chan} the channel from Bob to Alice.
\texttt{net\_send(pchan, m, reliable)} models sending a message \texttt{m} to
\texttt{\^{}pchan}, where \texttt{pchan} is either \texttt{\&s\_chan}
or \texttt{\&r\_chan}.
The method places either \texttt{m} (to model a successful send)
or \texttt{()} (to model loss) in \texttt{\^{}pchan}.
The use of the \texttt{reliable} flag will be explained later.
\texttt{net\_recv(pchan)} models checking \texttt{\^{}pchan} for
the next message.
